---
title: "Homework Assignment 1"
author: "Tien Huynh (3408549)"
date: "`r format(Sys.Date(), '%B %d, %Y')`"
output: pdf_document
---


```{r setup, echo=FALSE}
library(knitr)
# set global chunk options: images will be 7x5 inches
knitr::opts_chunk$set(fig.width=7, fig.height=5)
options(digits = 4)


## indents are for indenting r code as formatted text
## They may need to be adjusted depending on your OS
# if your output looks odd, increase or decrease indent
indent1 = '    '
indent2 = '        '
indent3 = '            '
```

1. The dataset cov.data contains number of confirmed COVID-19 cases (in thousand people), and population (in thousand people) of 9 counties in California (as of 1/18/2021).

```{r problem-1-info, include=FALSE}
cov.data <- data.frame(row.names = c("San Bernardino County",
                                      "Riverside County",
                                      "Orange County",
                                      "San Diego County",
                                      "Santa Clara County",
                                      "Kern County",
                                      "Sacramento County",
                                      "Fresno County",
                                      "Alameda County"))
cov.data$population <- c(2149, 2411, 3168, 3316, 1927, 887, 1525, 985, 1657)
cov.data$confirmed <- c(251, 239, 223, 212, 93, 85, 79, 81, 67)
```

a) 
The predictor variable in this problem is the county population. The response variable is the number of confirmed COVID-19 cases of a county.

b)  
```{r beta-calculation}
colof1 <- matrix(1, nrow=9, ncol=1)
population <- matrix(cov.data$population)
X <- cbind(colof1, population)
Y <- matrix(cov.data$confirmed)
beta_hat <- solve(crossprod(X,X), crossprod(X,Y))
rownames(beta_hat) <- c("(Intercept)", "population")
beta_hat
```
$\hat\beta_0 = 2.85395$
$\hat\beta_1 = 0.07236$

c)  
$\hat\beta_0$ represents the estimated number of confirmed COVID-19 cases of a county when the population of that county is 0. It is also the y-intercept of the linear regression model for the dataset when it is graphed. $\hat\beta_1$ represents the estimated amount of change in the number of confirmed COVID-19 cases of a county when the population of that county is increased by 1 unit. In other words, it is the slope of this specific simple linear regression model.

d) 
The simple linear regression model in this problem is $y_i = 2.85395 + 0.07236 x_i + \varepsilon_i$.

e)
```{r}
pop = 1000000/1000
calc = 2.85395+(0.07236*pop)
predicted = calc * 1000
predicted
```
A county with a population of 1,000,000 is predicted to have 75,214 confirmed COVID-19 cases.

2. The dataset trees contains measurements of Girth (actually, tree diameter) in inches, Height in feet, and Volume of timber (in cubic feet) of a sample of 31 felled black cherry trees. The following commands can be used to read the data into R. The dataset can be accessed under the name
trees.

```{r problem-2-info, include=FALSE}
# the dataset "trees" is contained in the R package "datasets"
require(datasets)
trees
```

a)
```{r tree-description}
?trees
```

The dataset trees has 31 observations (rows) and 3 variables (columns). These variables are Girth, Height, and Volume.

b)
```{r pairs-function}
pairs(formula = log(Volume) ~ log(Girth) + log(Height), data=trees)
```

c)
```{r correlation-matrix}
cormatrix <- cor(log(trees))
cormatrix # correlation matrix for the three (logged) variables
```

d)
```{r check-missing-value}
sum(is.na(trees))
```

There are no missing values.

e)
```{r lm-function}
mod <- lm(formula = log(Volume) ~ log(Girth) + log(Height), data = trees)
summary(mod)
```

Therefore, the multiple regression model in this problem is $log Volume_i = -6.632 + (1.983)log Girth_i + (1.117)logHeight_i + \varepsilon_i$.

f)
```{r design-matrix-check}
colof1 <- matrix(1, nrow=31, ncol=1)
logGirth <- matrix(log(trees$Girth))
logHeight <- matrix(log(trees$Height))
X <- cbind(colof1, logGirth, logHeight) # design matrix
Y <- matrix(log(trees$Volume))
beta_hat <- solve(crossprod(X,X), crossprod(X,Y))
rownames(beta_hat) <- c("(Intercept)", "log(Girth)", "log(Height)")
beta_hat
```

g)
```{r computations}
predict(mod) # predicted response values from fitted regression model
residuals(mod) # residuals
matrix <- (Y-(X%*%beta_hat))
numerator <- crossprod(matrix,matrix)
denominator <- mod$df.residual
eev <- numerator/denominator
eev # estimate of the error variance
```

3. (Answer located in a separate PDF.)