---
title: "Homework 2"
author: "Tien Huynh (3408549); PSTAT 126, Winter 2021"
date: "__Due date: Feburary 10, 2021 at 23:59 PT__"
graphics: yes
output: pdf_document
---

__Note:__ Please show all the procedures of your analysis, and prepare the homework solution using RMarkdown. All code should be well documented. A RMarkdown homework template is available on GauchoSpace. Homework should be submitted on GauchoSpace.

You should write up your homework solution on your own. In particular, do not share your homework RMarkdown file with other student.
 
---------------------------------

Q1. This question uses the _fat_ dataset available in the Homework 2 folder on GauchoSpace. The following command can be used to read the data into R. Make sure the "fat.csv" file is in the same folder as your R/Rmd file.

```{r load, message=F, warning=F, results="hide"}
fat <- read.csv(file = "fat.csv")
```

The dataset _fat_ contains measurements for 252 men on the following variables:

- brozek: Percent body fat 
- age: Age(yrs)
- weight: Weight(lbs)
- height: Height(inches)
- neck: Neck circumference(cm)
- chest: Chest circumference(cm)
- abdom: Abdomen circumference (cm)
- hip: Hip circumference (cm)

Our goal is to study how _brozek_ is related to all other 7 variables.

  (a). Use the _lm_ function in R to fit the MLR model with
  _brozek_ as the response and the other 7 variables as
  predictors. Display the summary output.
  
```{r problem-1-info}
mod <- lm(formula = brozek ~ ., data = fat)
summary(mod)
```
  
  
  (b). Which predictor variables are statistically significant under the significance threshold value of 0.05? 

The Abdomen Circumference (abdom) and Neck Circumference (neck) are the predictor variables that are statistically significant under the significance threshold value of 0.05. 

  (c). What proportion of the total variation in the response is
  explained by the predictors?
  
```{r r-squared}
rsquared <- summary(mod)$r.squared
percent <- rsquared * 100
percent
```
  
  As shown above, 72.71498% of the total variation in the response is explained by the predictors.
  
  (d). What is the p-value for the global F-test. Do the 7 predictor variables explain a significant proportion of the variation in the response? 
  
```{r global-f-test}
mod_M <- lm(brozek ~ ., fat)
mod_m <- lm(brozek ~ 1, fat)
anova(mod_m, mod_M)
```
  
The p-value for the global F-test is less than $2.2*10^{-16}$ (0.00000000000000022). Therefore, the 7 predictor variables explain a significant proportion of the variation in the response.
  
  (e). Consider testing the null hypothesis $H_0: \beta_{neck} = 0$, where $\beta_{neck}$ is the coefficient corresponding to _neck_ in the MLR model. Use the _t value_ available in the summary output to compute the p-value associated with this test, and verify that the p-value you get is identical to the p-value provided in the summary output. 
  
```{r t-test}
# t-test

n <- dim(fat)[1] # number of observations
p <- 7 # number of predictor variables
t_value <- coefficients(summary(mod))[5,3] # t-value
p_value <- pt(q = t_value, df = n-p-1) * 2
p_value
p_value == coefficients(summary(mod))[5,4]
```

  
  (f). Suppose we are interested in either _weight_ or _height_ had any relation to the _brozek_. This corresponds to a null hypothesis of $H_0: \beta_{weight} = \beta_{height} = 0$. Construct a F-test, report the corresponding p-value, and your conclusion.
  
```{r f-test}
# f-test

mod_M1 <- lm(brozek ~ ., fat)
mod_m1 <- lm(brozek ~ age + neck + chest + abdom + hip, fat)
anova(mod_m1, mod_M1)
```
  The corresponding p-value is 0.009257. Given that the significance threshold value is 0.05, the null hypothesis is rejected. Therefore, $\beta_{weight}$ and/or $\beta_{height}$ is/are not equal to zero.
  
  (g). Use the summary output to construct a 99\% confidence interval for $\beta_{abdom}$. What is the interpretation of this confidence interval?
  
```{r confidence-interval}
library(tibble)
abdom_std_error <- coef(summary(mod))[7,2]
adbom_beta_hat <- coefficients(mod)[7]
t_pct <- qt(p = 0.995, df = n-p-1)
CI_99 <- tibble(lower_bound = adbom_beta_hat - t_pct*abdom_std_error,
                upper_bound = adbom_beta_hat + t_pct*abdom_std_error)
CI_99
```
  The 99% confidence interval for $\beta_{abdom}$ is (0.7033266, 1.111842). This means that the interval ranging from 0.7033266 to 1.111842 cm is expected to contain the true value of $\beta_{abdom}$ 99% of the time.
  
  (h). What is the predicted _brozek_ for a person with the following information:
  
  - age = 40
  - weight = 170
  - height = 70
  - neck = 38
  - chest = 100
  - abdom = 91
  - hip = 99
  
```{r predict-brozek}
x0 <- data.frame(age = 40, weight = 170, height = 70, neck = 38, chest = 100, abdom = 91, hip = 99)
predicted <- predict(mod, newdata = x0, interval = "prediction")
predicted[1]
```
  The predicted _brozek_ for a person of those characteristics is 18.23699% body fat.
  
  (i). What is the 95\% prediction interval for the observation in part (h)? What is the interpretation of this prediction interval?
  
```{r}
x0 <- data.frame(age = 40, weight = 170, height = 70, neck = 38, chest = 100, abdom = 91, hip = 99)
predicted <- predict(mod, newdata = x0, interval = "prediction", level = .95)
predicted[2:3]
```
The 95\% prediction interval for the observation in part (h) is (10.10949, 26.36448). This means we are 95\% confident that the future response will fall within the interval 10.10949 and 26.36448.

----------------------------------------------

Q2. Consider the multiple linear regression model:
$$
\mathbf{y} = \mathbf{X} \beta + \varepsilon, \qquad \text{where} \qquad \varepsilon \sim N(\mathbf{0}, \sigma^2 \mathbf{I}_n).
$$
Show that the covariance matrix of the (ordinary) least squares estimate has the following form:
$$
\mathrm{Cov}(\hat{\beta}) =  \sigma^2 (\mathbf{X}^T \mathbf{X})^{-1}.
$$


Hint: For a generic random vector $z \in \mathbb{R}^p$ and a fixed matrix $\mathbf{A} \in \mathbb{R}^{m \times p}$, we have
$$
\mathrm{Cov}(\mathbf{A} z) = \mathbf{A} \mathrm{Cov}(z) \mathbf{A}^T.
$$

(Answer located in a separate PDF.)